---
title: "Index"
author: "Thomas Adams and Elena Spielmann"
format: html
editor: visual
editor_options: 
  chunk_output_type: console

---

# Global and National Security

## What it means and where the United States fits in

National security is the security and defense of a sovereign state, including its citizens, economy, and institutions, which is regarded as a duty of government. While it was originally conceived as protection against military attack, national security is widely understood to include also non-military dimensions, including the security from terrorism, minimization of crime, economic security, energy security, environmental security, food security, and cyber-security. Similarly, national security risks include, in addition to the actions of other nation states, action by violent non-state actors, by narcotic cartels, and by multinational corporations, and also the effects of natural disasters.

## A By-Country Look at Active-Duty Military Versus Total Military per Capita Globally

## Countries with larger total and active duty forces are not the same per captia... and what it means. 

```{r}

library(readxl)
library(ggplot2)
library(dplyr)

globalmilitary <- read_excel("C:/Users/tj/OneDrive/Desktop/datascience/week04/assignment04/globalmilitary.xlsx")


# Rename "People's Republic of China" to "China"
globalmilitary_filt <- globalmilitary %>% 
  mutate(Country = ifelse(Country == "People's Republic of China", "China", Country))

# Define a vector of the countries to keep
countries_to_keep <- c("Russia", "Vietnam", "India", "Switzerland", "United States of America", "Iran", "China", "North Korea", "Ukraine", "Pakistan", "Thailand", "Cuba", "Brazil", "Zimbabwe")

# Subset the data frame to only include the countries in the vector
globalmilitary_filt <- subset(globalmilitary_filt, Country %in% countries_to_keep)


#Get rid of scientific notation
options(scipen=999)

#Bubble plot reflecting total military versus active duty military per capita.

ggplot(data = globalmilitary_filt,
mapping = aes(x = active_military,
y = total,
size = per1000_capita_total)) +
geom_point(shape = 21, alpha = 0.5, color = "white", fill = "red") +
scale_size(name = "Per 1,000 Capita Total",
range = c(1, 10), breaks = c(20, 150, 300),
labels = c("20", "150", "300"),
limits = c(0, NA)) +
geom_text(aes(label = Country), showlegend = F, size = 2) +
scale_y_continuous(limits = c(0, 8100000, breaks = seq(0, 8100000, 500000))) +
scale_x_continuous(limits = c(0, 2200000, breaks = seq(0, 2200000, 200000))) +
theme_minimal() +
labs(x = "Active Duty Military",
y = "Total Military", 
title = "Active-Duty Military Versus Total Military per Capita Globally",
subtitle = "Countries with larger total and active duty forces are not the same per captia.",
caption = "Source: The Military Balance (2017) | International Institute of Strategic Studies") +
   theme(axis.text.y = element_text(angle = 0),
        axis.title.y = element_text(angle = 0, vjust = 1))

#NOTES: need to hide warning labels

```

Total military is comprised of active military, reserve military and paramilitary. Countries with low numbers of both total military and active military tend to be countries that either have peaceful diplomatic agendas (ie. Switzerland) or countries that simply cannot afford to maintain a large military force (ie. Zimbabwe). Conversely, countries with less peaceful diplomatic agendas and greater financial resources have higher numbers of total military and active military (ie. China)

Viewing each of these countries total and active military numbers through a "per capita" lens is additional useful in identifying how high national security is prioritized by country and the financial and population resources available to it. For example, North Korea has the highest number of total and active duty military number per captia, but this does not reflect it's economic power. This is also true for countries such as the United States and China, which have large total populations and total and active military numbers, but these numbers are not high in comparison to North Korea per captia. 

## A Look at the United States Allies and Advarsaries Military Spending Across time

## United States and allies more transparant about military spending compared to near peer threats

```{r}

library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)

military_spending <- read_csv("C:/Users/tj/OneDrive/Desktop/datascience/week04/assignment04/military_spending.csv")

#Re-code to select needed variables

near_peer <- c("China", "European Union", "Iran", "Russian Federation",
"United States")

near_peer.spend <- filter(military_spending,
(Name %in% near_peer))

#Re-code data from wide to long

select_mil_spend <- pivot_longer(near_peer.spend,
names_to = "year",
values_to = "spending", 5:50)

#Rename Russia
select_mil_spend <- select_mil_spend %>% 
  mutate(Name = if_else(Name == "Russian Federation", "Russia", Name))

#Factorize to specify order of legend

select_mil_spend$Name <- factor(select_mil_spend$Name, 
                                levels = c("United States", "European Union", 
                                           "China", "Russia", "Iran"))

#Convert military spending totals to billions

select_mil_spend$spending <- select_mil_spend$spending / 1e9

#Time series line plot

ggplot(data = select_mil_spend,
mapping = aes(x = as.numeric(year), y = spending)) +
geom_line(mapping = aes(color = Name)) +
    scale_color_manual(values = c("United States" = "green",
                                  "European Union" = "blue",
                                    "China" = "red", 
                                  "Russia" = "black",
                                "Iran" = "orange"
                                )) +
  geom_vline(xintercept = 1993, linetype = "dashed", color = "gray") +
  geom_text(x = 1993, y = 400, label = "Russia\n(1993)", 
            hjust = -0.1, vjust = -0.5, color = "gray") +
  geom_vline(xintercept = 1989, linetype = "dashed", color = "gray") +
  geom_text(x = 1989, y = 450, label = "China\n(1989)", 
            hjust = -0.1, vjust = -0.5, color = "gray") +
xlab("") +
ylab("") +
labs(title = "U.S. and Allies Military Spending Compared to Advasaries \nMilitary Spending Across Time",
subtitle = "United States and allies more transparant about military spending \ncompared to near peer threats",
caption = "Source: The World Bank Database",
x ="Year",
y = " Military Spending \n(in Billions)",
color = "Country") +
theme(text = element_text(size = 10)) +
theme_minimal() +
     theme(axis.text.y = element_text(angle = 0),
        axis.title.y = element_text(angle = 0, vjust = 1))

```

Looking at military spending over time highlights the lack of transparency regarding military spending between Russia and China compared to the United States and our alliances (Europe). Russia and China are considered “near-peer” threats to the United States because of their increasing military power and strength regarding weapons, technology, and force. Iran is included because, while it is a threat to the United States, it is not considered “near-peer.” Iran is a useful gauging point to compare the reported spending of China and Russia because it is highly unlikely that China and Russia are actually spending close to the same amount of money on their military as Iran. China began reporting military spending in 1989 and Russian began reporting in 1993. Both countries are likely still not fully transparent with current reported spending amounts. Monitoring military spending is an important aspect to national and global security and the balance of power. It is likely that threats to the United States, such as China and Russia, are growing their military power, but through unclear means that should perhaps be further investigated.

```{r}
#| output: false
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(tigris)
library(remotes)
library(rgeoboundaries)
library(sf)
library(tinytex)
```
```{r}
#| warning: false
terrordata <- read_xlsx("C:/Users/tj/OneDrive/Desktop/datascience/week04/assignment04/globalterror.xlsx")
```
```{r}
#| output: false
terrordata <- terrordata[!is.na(terrordata$longitude), ]
terrorsf <- st_as_sf(terrordata, coords = c("longitude", "latitude"))
terrorsf <- terrorsf %>% filter(country == 217)
united.states <- geoboundaries(country = "United States of America", adm_lvl = "adm1", type = "SSCGS")
united.states <- united.states %>% filter(shapeName != "Guam") %>% filter(shapeName != "Commonwealth of the Northern Mariana Islands") %>% filter(shapeName != "American Samoa") %>% filter(shapeName != "Alaska")
st_set_crs(united.states, value = 4326)
terrorsf <- st_set_crs(terrorsf, value = 4326)
```
```{r}
#Sorting by number of fatalities in an attack
k_terrorsf <- terrorsf %>%
  filter(., nkill > 0) %>%
  arrange(., nkill)

  ggplot(data = united.states) +
  scale_color_gradient2(
    low = "green",
    mid = "yellow",
    high = "red",
    midpoint = 750
  ) +
  geom_sf(
    data = united.states
  ) +
  geom_sf(
    data = k_terrorsf,
    aes(color=nkill, size = nkill, alpha = .02) 
  ) + 
  labs(title = "Fatal Terror Attacks In the US", subtitle = "Scaled according to Fatalities", color = "Deaths", size = "Deaths", caption = "Source: START Global Terrorism Database") +
  theme_void()
  ```